---
output:
  word_document: default
  html_document: default
---
# Module 3 - Assignment 1
# Subramanian, Subbu
# Data Visulaization

Will be using the datasets containing candy rankings and production that can be found on the Canvas course website.

```{r Loading Tidyverse and data}

library(tidyverse)
candy_data<-read_csv("~/Documents/MBA/Programming for Analytics/Module 3/Assignment 1/candy_data.csv")
candy_production <- read_csv("~/Documents/MBA/Programming for Analytics/Module 3/Assignment 1/candy_production.csv")
```

# Visualization with Scatterplots (geom_point)

```{r Scatterplot}
library(tidyverse)
ggplot(data=candy_data,aes(x=sugarpercent,y=pricepercent,label=competitorname)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_text(check_overlap = T,
              vjust = "bottom", 
              nudge_y = 0.01, 
              angle = 30,size = 2) +
  labs(title = "Sugar by Price Scatterplot", 
         x = "Sugar content (percentile)", 
         y = "Price (percentile)")
```

On analysis of the scattered plot of the candy data, "Skittles Original" has the most sugarpercent and lowest pricepercent.

# Line Chart of Candy Production

Will be using the datasets containing candy production dataset which will display a specific date and how production during that month is compared to 2012

```{r Line chart of candy production}
ggplot(data=candy_production,aes(x=observation_date,y=IPG3113N)) +
  geom_line() +
   labs(title = "Monthly Candy Production(US)", 
         x = "Year", 
         y = "As percent of 2012 production")


```

# Bar Chart of Ingredients

```{r}
library(tidyverse)
candyFeatures<-candy_data %>% select(2:10) 
candyFeatures[]<- lapply(candyFeatures, as.logical)
ggplot(data=candyFeatures,aes(x=chocolate)) +
  geom_bar()
```

