---
output:
  word_document: default
  html_document: default
---

# Module 4 - Assignment 1
# Subramanian, Subbu
# Data Transformation

```{r Data Load}
library(tidyverse)
state_income <- read_csv("state_income.csv")
```

## State Incomes:

I will be creating a subset of data from the file and will be using the below variables for the analysis,

1.) State_Name
2.) State_ab
3.) County
4.) City
5.) Type
6.) ALand
7.) Mean
8.) Median
9.) Stdev


```{r Select}

State_income2<-select(state_income,State_Name,State_ab,County,City,Type,ALand,Mean,Median,Stdev)
select(State_income2,State_ab,everything())
head(State_income2,10)

```

```{r Rename}

State_income2<-rename(State_income2,SquareArea=ALand,MeanIncome=Mean,IncomeMedian=Median,IncomeStDev=Stdev)
head(State_income2,10)
```

```{r Filter}

NC_income<-filter(State_income2,State_Name=="North Carolina")
head(NC_income,10)
```

## NC Incomes:

I will be using the NC_income dataset to create summaries of the incomes within North Carolina including summaries by county, city and type.

```{r Arrange}

NC_income<-arrange(NC_income,County)
head(NC_income,10)

```

```{r Summarise}

summary1 <- group_by(NC_income,County)
summary1 <- summarise (summary1, mean= mean(MeanIncome))
summary2 <- NC_income %>%
group_by(City) %>%
summarise (mean = mean(MeanIncome))

```


Summary 1 Dataset provides the mean income for every county in North Carolina

Summary 2 Dataset provides the mean income for every city in North Carolina

```{r Summarize Continued}
summary3 <- NC_income %>%
group_by(Type) %>%
summarise (mean = mean(MeanIncome))
head(summary3,10)
```


## Income Visualization

```{r Plot}


ggplot(data=summary1,aes(x=County,y=mean,label=County)) +
geom_point() +
   labs(title = "County Income Data", 
         y = "Income") +
theme(axis.text.x = element_text(angle = 90,vjust =0.5, hjust=1)) 
 

```

* County with the largest average Income : Wake County
* County with the lowest average income : Stokes County
* County with the second lowest income : Surry County


```{r Average State income}


AvgStateIncome <- group_by(State_income2,State_ab)
AvgStateIncome <- summarise (AvgStateIncome, mean= mean(MeanIncome))
AvgStateIncome <-filter(AvgStateIncome,grepl("N",State_ab))
ggplot(data=AvgStateIncome,aes(x=State_ab,y=mean)) +
geom_point() +
   labs(title = "State Income Data", 
         y = "Income") +
theme(axis.text.x = element_text(angle = 90,vjust =0.5, hjust=1)) 
  



```

* State with the largest average Income : NJ
* State with the lowest average income : TN