---
output:
  word_document: default
  html_document: default
---


# Module 5 - Assignment 1
# Subramanian, Subbu
# Data Wrangling

```{r Importing Data}
library(tidyverse)
library(readxl)
UN_migrant <- read_excel("~/Documents/MBA/Programming for Analytics/Module 5/Assignment 1/Mod5Assignment1/UN_migrant.xlsx", 
    sheet = "Table 6", col_types = c("numeric", 
        "text", "text", "numeric", "text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric"), skip = 15)

```

## Part 2 – Cleaning Data with dplyr

```{r Renaming Columns}
UN_migrant<-rename(UN_migrant,"Country"="...2","Country_Code"="...4","Type"="...5","1990"="1990...6","1995"="1995...7","2000"="2000...8","2005"="2005...9","2010"="2010...10","2015"="2015...11")

Migration<-select(UN_migrant,"Country","Country_Code","Type","1990","1995","2000","2005","2010","2015")
as_tibble(Migration)

```

# Part 3 – Creating tidy data using tidyr

```{r Creating Tidy Data}
Migration2 <- pivot_longer(Migration,c('1990','1995','2000','2005','2010','2015'),names_to="year",values_to="cases")
Migration2$year<-as.integer(Migration2$year)
head(Migration2,6)

```

## Part 4 - Research Questions

```{r Visualization}

RegionalMigration<-Migration2 %>%
  select(Country,Country_Code,Type,year,cases) %>%
  filter(Country %in% c("Africa", "Asia", "Europe", "Oceania", "Northern America", "South America")) 

Americas<-Migration2 %>%
  select(Country,Country_Code,Type,year,cases) %>%
  filter(Country %in% c("Central America", "South America", "Northern America")) 

```

## Wordwide Migration based on Regions

```{r Plotting for World}

ggplot(data=RegionalMigration,aes(x=year,y=cases,color=Country)) +
  geom_point() +
  labs(title = "Worldwide Migration", x = "Year", y="Migrants")

ggplot(data=RegionalMigration,aes(x=year,y=cases,fill=Country)) +
  geom_col() +
  labs(title = "Worldwide Migration", x = "Year", y="Migrants") +
  scale_fill_discrete(name="Region of World")
```


* a. Which region in the world had the highest number of migrants in the year 2005? Asia
* b. Over the years, which region consistently has the most migrants every 5 year span? Which has the second most? First = Asia, Second = Africa
* c. What region has seen the fewest migrants over the years? Oceania
* d. Which plot was most useful in answering these questions and why? Bar/col, because it was visually easier to figure out the regions, whereas with scattter plot the size of the points were too small to see and determine the correct answer.

## Americas Migration by Region

```{r Plotting for Americas}

ggplot(data=Americas,aes(x=year,y=cases,color=Country)) +
  geom_point() +
  labs(title = "Worldwide Migration", x = "Year", y="Migrants")

ggplot(data=Americas,aes(x=year,y=cases,fill=Country)) +
  geom_col() +
  labs(title = "Worldwide Migration", x = "Year", y="Migrants") +
  scale_fill_discrete(name="Americas Region")
```

* a. In 1990, which region had the largest number of migrants for the Americas? Central America
* b. Has this region continued to grow since 1990? No
* c. What trends do you notice happening in the Americas over the years? Migrants to Central America has subsided
* d. Specifically, has Northern America increased or decreased over the years? Increased over the years
* e. Which plot was most useful in answering these questions and why? Bar/col, because it was visually easier to figure out the regions, whereas with scattter plot the size of the points were too small to see and determine the correct answer.



