---
output:
  word_document: default
  html_document: default
---


# MIS503 – Final Project
# Subramanian, Subbu
# Zillow Home Value Index Analysis

```{r Loading Packages and Data}
library(tidyverse)
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
```

## Wake County Home Sales

```{r Wake County}

WakeCountySales<- SingleFamilyResidenceSales %>%
  select(RegionName,State,CountyName,Metro,`1996-05`,`1997-05`,`1998-05`,`1999-05`,`2000-05`,`2001-05`,`2002-05`,`2003-05`,`2004-05`,`2005-05`,`2006-05`,`2007-05`,`2008-05`,`2009-05`,`2010-05`,`2011-05`,`2012-05`,`2013-05`,`2014-05`,`2015-05`,`2016-05`,`2017-05`,`2018-05`) %>%
  filter(State %in% c("NC"), CountyName %in% c("Wake County")) 

WakeCountySales<- rename(WakeCountySales,"1996"="1996-05","1997"="1997-05","1998"="1998-05","1999"="1999-05","2000"="2000-05","2001"="2001-05","2002"="2002-05","2003"="2003-05","2004"="2004-05","2005"="2005-05","2006"="2006-05","2007"="2007-05","2008"="2008-05","2009"="2009-05","2010"="2010-05","2011"="2011-05","2012"="2012-05","2013"="2013-05","2014"="2014-05","2015"="2015-05","2016"="2016-05","2017"="2017-05","2018"="2018-05")

as_tibble(WakeCountySales)

WakeCountySales<- pivot_longer(WakeCountySales,c('1996','1997','1998','1999','2000','2001','2002','2003','2004','2005','2006','2007','2008','2009','2010','2011','2012','2013','2014','2015','2016','2017','2018'),names_to="YR",values_to="ZHVI")

ggplot(data=WakeCountySales,aes(x=YR,y=ZHVI,color=RegionName)) +
  geom_point() +
  labs(title = "Wake County Home Sales", x = "Year", y="Home Values (in$)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)

ggplot(data=WakeCountySales,aes(x=YR,y=ZHVI,color=RegionName)) +
  geom_col() +
  labs(title = "Wake County Home Sales", x = "Year", y="Home Values (in$)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)



```

* What have been the overall trends in Wake County Home Values? Values have been consistently rising except for a small dip in the years 1998, 2009, 2010, 2011 and 2012. 
* There were dips in home values in the past 20 years. What years did these occur? 1998, 2009, 2010, 2011 and 2012. The reason for dip between 2009 to 20102 could be due to the recession that happened during 2008 to 2009.
* Based on the analysis, where would be the least expensive area to purchase home? Most expensive area? Least expensive area is Zebulon, most expensive are is Raleigh.
* Are any area home values trending down? Is there one area that stands out compared to others? New Hill seems to have dropped a bit between 2017 to 2018


## NC Rental Market

```{r NC Rental}

Rentals<- SingleFamilyResidenceRental %>%
  select(RegionName,State,`2010-11`,`2011-11`,`2012-11`,`2013-11`,`2014-11`,`2015-11`,`2016-11`,`2017-11`,`2018-10`) %>%
  filter(State %in% c("NC"), RegionName %in% c("Asheville", "Charlotte", "Durham", "Fayetteville", "Raleigh", "Wilmington")) %>%
  rename("2010"="2010-11","2011"="2011-11","2012"="2012-11","2013"="2013-11","2014"="2014-11","2015"="2015-11","2016"="2016-11","2017"="2017-11","2018"="2018-10") %>%
  pivot_longer(c('2010','2011','2012','2013','2014','2015','2016','2017','2018'),names_to="YR",values_to="ZHVI")

as_tibble(Rentals)

ggplot(data=Rentals,aes(x=YR,y=ZHVI,color=RegionName)) +
  geom_point() +
  labs(title = "Rental Costs in NC Cities", x = "Year", y="Average Rental Cost (in$)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average Rental Cost (in$)", labels = scales::comma)

ggplot(data=Rentals,aes(x=YR,y=ZHVI,color=RegionName)) +
  geom_col() +
  labs(title = "Rental Costs in NC Cities", x = "Year", y="Average Rental Cost (in$)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average Rental Cost (in$)", labels = scales::comma)


```

* What has been the overall trend in the rental market around the state? Are there any cities that have not followed this trend? Rental has been on the increase overall, except for Fayetteville where it is almost around the same pricing and even had a dip in 2016.
* Where is the most expensive city to rent in? Least expensive? Asheville is the most expensive city rent in whereas Fayetteville is the lease expensive.
* You are trying decide between Wilmington and Asheville. Which market has the lowest rent? Wilmington.


## Home Values in Select Rental Markets

```{r Home Values in Select Rental Markets}

NCHomeSales<- SingleFamilyResidenceSales %>%
  select(RegionName,State,Metro,`1996-05`,`1997-05`,`1998-05`,`1999-05`,`2000-05`,`2001-05`,`2002-05`,`2003-05`,`2004-05`,`2005-05`,`2006-05`,`2007-05`,`2008-05`,`2009-05`,`2010-05`,`2011-05`,`2012-05`,`2013-05`,`2014-05`,`2015-05`,`2016-05`,`2017-05`,`2018-05`) %>%
  filter(State %in% c("NC"), Metro %in% c("Asheville", "Charlotte-Concord-Gastonia", "Raleigh", "Wilmington")) %>%
  rename("1996"="1996-05","1997"="1997-05","1998"="1998-05","1999"="1999-05","2000"="2000-05","2001"="2001-05","2002"="2002-05","2003"="2003-05","2004"="2004-05","2005"="2005-05","2006"="2006-05","2007"="2007-05","2008"="2008-05","2009"="2009-05","2010"="2010-05","2011"="2011-05","2012"="2012-05","2013"="2013-05","2014"="2014-05","2015"="2015-05","2016"="2016-05","2017"="2017-05","2018"="2018-05") %>%
  pivot_longer(c('1996','1997','1998','1999','2000','2001','2002','2003','2004','2005','2006','2007','2008','2009','2010','2011','2012','2013','2014','2015','2016','2017','2018'),names_to="YR",values_to="ZHVI") %>%
  group_by(Metro)

as_tibble(NCHomeSales)

ggplot(data=NCHomeSales,aes(x=Metro,y=ZHVI)) +
  geom_violin() +
  geom_boxplot(width=0.1)
  



```

* According to the results, which market has the lowest median price (represented as horizontal bar in box plot)? Raleigh
* The violin plot will show density meaning the wider the plot is, the more observations occur within that area. Which market has the most density around the median value of homes? Charlotte-Concord-Gastonia


## Relocation Home Value Comparison

```{r Relocation Home Value Comparison}

NationalHomeSales<- SingleFamilyResidenceSales %>%
  select(RegionName,State,Metro,`1996-05`,`1997-05`,`1998-05`,`1999-05`,`2000-05`,`2001-05`,`2002-05`,`2003-05`,`2004-05`,`2005-05`,`2006-05`,`2007-05`,`2008-05`,`2009-05`,`2010-05`,`2011-05`,`2012-05`,`2013-05`,`2014-05`,`2015-05`,`2016-05`,`2017-05`,`2018-05`) %>%
  filter(State %in% c("IL", "CO", "TX", "NY"), RegionName %in% c("Chicago", "Denver", "Houston", "New York")) %>%
  rename("1996"="1996-05","1997"="1997-05","1998"="1998-05","1999"="1999-05","2000"="2000-05","2001"="2001-05","2002"="2002-05","2003"="2003-05","2004"="2004-05","2005"="2005-05","2006"="2006-05","2007"="2007-05","2008"="2008-05","2009"="2009-05","2010"="2010-05","2011"="2011-05","2012"="2012-05","2013"="2013-05","2014"="2014-05","2015"="2015-05","2016"="2016-05","2017"="2017-05","2018"="2018-05") %>%
  pivot_longer(c('1996','1997','1998','1999','2000','2001','2002','2003','2004','2005','2006','2007','2008','2009','2010','2011','2012','2013','2014','2015','2016','2017','2018'),names_to="YR",values_to="ZHVI") %>%
  group_by(Metro)

ggplot(data=NationalHomeSales,aes(x=YR,y=ZHVI)) +
  geom_point() +
  facet_wrap(~Metro) +
  labs(title = "City Home Value Comparison", x = "Year", y="Home Values (in$)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size=5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)
  

```


* Based on your analysis, which city’s housing is most affordable? Least affordable? Houston
* Which cities saw the largest change in prices over the past 5 years? Which city has remained more consistent (i.e., no huge swings up or down in home values)? Largest Change = NewYork-Newark-Jersey City, Consistent = Houston-The Woodlands-Sugar Land
* During the market downturn in 2012, which cities were most impacted? Which cities have recovered? Most Impacted = Chicago-Napeville-Elgin, Recovered Cities = NewYork-Newark-Jersey City, Denver-Aurora-Lakewood

